A Tool for Visualizing Patterns of Spreadsheet Function Combinations

### 2016-04-20 doc/paper.tex
Paper syncronization.

	\section{Introduction} Business and research alike a lot to spreadsheets, the
	tabular interface that offers users a structured way to store and manipulate
	potentially huge bodies of data. Their allure is in their versatility: while
	the novice can still work without a deep knowledge of programming, they can
	learn to expedite their work by building programs out of several available
	functions \cite{nardi1990spreadsheet}. As such, it should not be surprising
	that when Scaffidi and colleagues estimated that over 50 million U.S. workers
	could be using them by 2012, with 25 million of them writing programs 
	\cite{scaffidi2005estimating}. \par Considering this ubiquity, it's crucial to
	understand how people are actually using spreadsheets. After all, failures in
	their practice can be ruinous, as demonstrated in EuSpRIG's many horror stories
	of spreadsheets gone wrong. In 2013, for example, they detail the influence of
	an economics paper on debt and national growth which shaped many political
	initiates around it, only to discover that the spreadsheets used in the data
	analysis contained an error that reversed the findings. The other stories,
	likewise, do not augur well if shoddy spreadsheet practices persist
	\footnote{http://www.eusprig.org/horror-stories.htm}. \par Fortunately, many
	spreadsheet researchers have assembled and released a number of spreadsheet
	corpora to inform work of how people actually use these tools. Collections like
	EUSES \cite{fisher2005euses}, FUSE \cite{barik2015fuse}, and the Enron corpus
	\cite{hermans2015enron} have already enabled fruitful work across the field.
	Hermans and colleagues, for example, used the EUSES corpus in 2011 to inform
	and evaluate their work on exploring code smells, or inelegant places in code
	that suggest the need for refactoring, between worksheets
	\cite{hermans2012detecting}. \par A tool, then, that empowers users of any
	intention to explore these datasets would be rife with potential. On a simple
	level, the tool could provide concrete statistics on which functions are used
	in practice most (or least) often for a given dataset. Beyond this, though, it
	could also inform future work on recommending functions through an
	understanding of this patterns, or it could guide API improvement by suggesting
	functions to add or prune based how people actually use them. Furthermore, if
	the tool maintains the connection between its broad patterns and the actual
	instances of formulae in workbooks, it could serve as a boon to educators not
	only by showing which functions and combinations are most practical to learn
	but also by offering a bounty of instructive, and real, examples \par This
	paper contributes a tool, informed by some of the aforementioned spreadsheet
	corpora, that seeks to accomplish these tasks by offering an interactive
	interface for exploring how people combine functions. It seeks to balance
	several needs, such as capturing these patterns in a broad enough overview
	while still be specific enough lead back to the precise places in spreadsheets
	where these patterns are realized. Along with the tool, then, is also a
	discussion of the various goals which the tool seeks to fulfill and the
	trade-offs that we negotiated at each point. All of this culminates in an
	evaluation of the tool through a study of possible cases to which the tool
	might apply.

	\section{Related Work} This tool comes from a line of spreadsheet visualization
	tools before it, each with a different focus. Some tools, such as Igarashi and
	colleague's fluid visualizations \cite{igarashi1998fluid}, seek to visualize
	the hidden formulas within a spreadsheet by imposing the dataflow graphs over
	the cells. Likewise, Clermont \cite{clermont2003scalable} (and Hipfl
	\cite{hipfl2008using}, who extended his work) explored various ways of
	visualizing groups of related cells through similar functions, neighbors, or
	references. Others focus on creating visualizations outside of the sheets:
	Hermans and colleagues address a spreadsheet programmer's information needs
	through their work on the tools GyroSAT \cite{hermans2011supporting} and Breviz
	\cite{hermans2011breviz} to make dataflow diagrams of individual spreadsheets.
	These, however, tend to focus on individual spreadsheets and not bodies of
	them. \par Nevertheless, other studies and tools focus more on evaluating the
	content of the cells therein. Badame, for example, implements a number of Excel
	formula refactorings and uses Euses in part to evaluate it
	\cite{badame2012refactoring}. Hermans, leveraging her work with the
	aforementioned dataflow diagrams, approached the same dataset to assess the
	prevalence of common inter-worksheet smells, such as feature envy and
	inappropriate intimacy \cite{hermans2012detecting}. Cunha and colleagues,
	likewise, constructed and refined a catalogue of smells and implemented the
	SmellSheet Detective tool to detect bad smells in chosen spreadsheets
	\cite{cunha2012towards}, which Abreu and colleagues later extended to detect
	faults. \cite{abreu2014smelling}.\par Implicit in this study is an exploration
	of how users interact with the Excel API. Previous research, such as the first
	study on the Enron spreadsheet corpus \cite{hermans2015enron}, on how users
	employ certain functions or commands. Researchers have applied such questions
	to other contexts as well, such as Murphy and colleague's study into how Java
	developers use Eclipse and what commands they use most often
	\cite{murphy2006java}.  Others focus on the documentation of the API itself:
	Robillard and DeLine, for example conducted a series of surveys and interviews
	to diagnose common problems with learning APIs, and they found that code
	examples, one of the focuses of this tool, is one of the most important aids to
	have in learning a new system \cite{robillard2011field}.

	\section{Approach/Methodology} \subsection{Walkthrough of Tool} Before
	discussing the minutiae of how the tool was developed, it will help first to
	give a basic example of how the tool can be used and how to interpret its
	symbols. One such example might be to find the answer to this question:
	\textit{ What kinds of functions do people use to define the condition in an IF
		function?} The final visualization is available in figure \ref{fig:fullpic};
	however, to understand the interaction, we will start rather from the
	beginning. \par When the user first approaches the tree for a given top-level
	function -- that is, a function nested within none other -- only a few nodes
	are visible, as shown in figure \ref{fig:startpic}. \par

@@

	The visualization so far comprises two types of node: the circle, which
	represents a discrete function in the formula and is size according to its
	relative frequency (the top-level node, by definition, will be the largest);
	and the numbered squares, which represent the positions of arguments within its
	parent function. From this, we can infer that, of the times it was observed, IF
	can have at most three arguments passed into it, which corresponds with its
	specifications in the
	API.\footnote{https://support.office.com/en-us/article/IF-function-69aed7c9-4e8a-4755-a9bc-aa8bbff73be2} \par Knowing that it is the first argument which contains the conditionals, we click on the square labeled "1" to explore. To save space, when more than 10 unique arguments have been observed in any position, the tool displays only the first ten, with an option to display the rest. The results are shown below:

@@

	As expected, we see that IF contains as its first argument a number of
	comparison operators, such as = and <=, and boolean-returning functions, like
	ISNA and AND, with simple equality being the most common and some use as ABS
	being the least seen of everything actually used. \par From here, we can
	further explore the common options among these functions. Clicking on the "="
	node will yield two arguments, it being a binary operator, and expanding each
	of them will peer into the range of common values of equality comparison, which
	Figure \ref{fig:fullpic} shows. \par For both sides of the equal sign, the
	operator has certain types of arguments which predominate over the others: on
	the left side is most often a reference to another cell; on the right, a string
	or number literal, which makes sense for the case of confirming a value in
	another cell before assigning this one. Furthermore, if the concept is
	difficult to imagine is practice, a tooltip accompanies each function node in
	the tree, providing a concrete example of a function that uses this structure
	and where it can be found. If this single instance is not enough, then the user
	also has the option to double-click the individual function node, which will
	open up a new tab with a table of many more examples from dataset.

@@

	\subsection{Design Goals} In visualizing the spreadsheet data, I outlined a few
	core goals for what the tool should accomplish: \begin{itemize} \item [1]
		\textbf{Draw an interactive interface to explore observed function
			combinations.} The combination space for all Excel functions is massive, let
		alone the space for observed formulas. Working from data with actual referents
		in practice, the tool must aid the user in navigating this space. \item [2]
		\textbf{Emphasize the quantitative patterns in formula construction.} The tool,
		accommodating datasets of a few spreadsheets to millions, should address the
		questions of how often the end users employed a certain function and where they
		used it. In this way, it must show precise metrics, such as frequency of  use
		and depth of function nesting, of the dataset it conveys. \item [3]
		\textbf{Promote a qualitative understanding of the patterns.} Lest these
		patterns remain abstract, the tool should supplement its observations with
		concrete instances of relevant formulas from the corpus. Ideally, it should
		even direct users to the exact cell in the spreadsheets where the formula was
		used, contextualizing the functions. \end{itemize} \par Likewise, to bound our
	scope, we outlined a few goals to specifically avoid accomplishing with this:
	\begin{itemize} \item [!1] \textbf{Do not try to directly explain what a
			function does.} Though the tool tries to foster tries understanding by linking
		pattern to example, it won't provide a precise description of what a function
		accomplishes. The tool's user must infer this. \item [!2] \textbf{Do not create
			new formulas.} This is essentially an exploratory tool, not a generative one.
		It should not produce any information other than new views of the original
		data. \item [!3] \textbf{Do not visualize individual formulas.} Though there is
		room to explore the place of a single function in the group, the tool must
		design the core visualization around the entire body of data, not the other way
		around. \end{itemize}

	\subsection{Design Decisions} Early in the process, we decided that the tree
	form would be a suitable fit to represent the parent-child or caller-callee
	relationships inherent in the data, given the composition of formulas as
	functions and their arguments (which could be yet more functions). By adapting
	this structure to accommodate a broad range of possibilities for nested
	functions, the branching factor depends on both the number of arguments in a
	function and the number of possible functions observed as an argument in a
	function. \par Guided by the goals in section A, we faced a number of decision
	points, a sample of which we describe below, before we arrived at the design
	shown above. \begin{itemize} \item \textit{Copied formulas}: Excel allows users
		to spread a formula over an area, repeating the same task in each cell with
		minor adjustments. Without checking for this, the analysis may not reveal the
		functions most commonly used together but rather the formulas most often
		applied to large areas. To combat this, we converted formulas from their native
		A1 format to the relative R1C1, in which copied formulas should be identical,
		and reduced the records to unique R1C1 formulas per sheet. \item
		\textit{Importance of depth}: When a function appears within another, should it
		be analyzed only as a nested function or would it also be valid to analyze the
		nested function on its own?  \par For example, in the pictured IF function, we
		found that people often use another IF statement as the second or third
		arguments of the top-level IF. If we only consider functions exactly where
		they're found embedded in the formula, then information about the same function
		-- IF, in this case -- will be scattered across different trees with no way to
		aggregate them. If we record every instance of a function by ignoring their
		context -- that is, including an IF embedded within a SUM function in the same
		node as the top-level IF -- then the tool will represent some functions in
		multiple nodes to capture every possible level of nesting. Both approaches have
		benefits and drawbacks, and so we included both. \item \textit{Pattern
			density}: How should the tool quantitatively order its elements: by a
		function's raw frequency or by the unity of patterns it leads to? For example,
		if SUM has for its first argument two possibilities, one which itself contains
		1000 unique argument possibilities with 1 occurrence each (high frequency, low
		pattern density) and another seen with 2 argument possibilities of 100
		occurrences each (low frequency, high pattern density), which would be more
		interesting to emphasize? The answer depends on the nuances of the questions,
		but for simplicity, I've shown the former. \item \textit{Non-functions}: How
		should the tool represent everything in the formula that isn't a function:
		numbers, string literals, errors, references, etc? Since these don't accept
		arguments, they will adorn the tree as leaves, and their precise content won't
		affect the functions around them as long as their types are known. As such, in
		the visualization, all of these nodes are replaced and aggregated under their
		types and represented as empty white nodes. \item \textit{Optional arguments}:
		How should the tool handle functions which accept a variable number of
		arguments? SUM, for example, can have anywhere from 1 to 255, and IF can accept
		either 2 or 3.
		\footnote{https://support.office.com/en-us/article/Excel-functions-by-category-5f91f4e9-7b42-46d2-9bd1-63f26a86c0eb} Without prior evidence, it's possible that, when people use optional arguments, they use different kinds of types and functions for arguments versus the cases when they ignore optional arguments. To account for this, we separate and analyze the different quantities of arguments observed for each function; the tool, however, uses as default all of the options collapsed into a single representation. \end{itemize}

@@

	\subsection{Bad Smells} Fowler's (?) seminal description of code smells
	underscored an important point of code quality: between perfect code and
	bug-crippled spaghetti,  there is a spectrum of code designs which, by
	themselves, are not faulty but nevertheless suggest problems and
	vulnerabilities in code design. Code smells, as we call them, soon commanded
	attention and earned wide-ranging applicability, eventually making it to the
	art of spreadsheet design. Since then, Hermans and others catalog the odors of
	spreadsheets, often built off Fowler's (?) original specifications to
	articulate the subtle problems in spreadsheet design. \par Collected, some of
	the intra-spreadsheet smells fall well within the purview of this visualization
	tool. Some such smells can be found in a figure [FIGURE]. \par The smells
	described above, then, entail a corresponding structure visible in the
	visualization: the deeper a branch runs, or the taller the range of arguments,
	the more it emits the stench of suboptimal design. \par To be sure, these rank
	occurrences must be leveraged with their prevalence across the dataset. One
	person's anomaly, though perhaps still interesting, will not be as impactful as
	a smell of thousands. \par Some examples are meaningful not for the space they
	take up but for how they clash with the expectations around a function. For
	example, Excel offers a number of functions that accept any number arguments
	given to them, up to a resource-defined limit. SUM and CONCATENATE are two such
	functions: adding more arguments simply adds or appends that element in the
	same way as every previous argument. Strange cases were found, then, in the
	cases where these functions were used with only one argument. \par Hermans, in
	her paper on the nuances of VLOOKUP and its common misuses, explore the problem
	of the final, optional argument. A tool there helped, but how might a
	visualization such as this illuminated the same?

@@

	\section{User Study} Midway through production, we conducted a brief,
	four-participant user study in order to gauge how well we were capturing the
	initial design goals. The study was, in general, a flexible and exploratory
	affair: rather than giving the users any specific tasks or pointed questions,
	we asked instead that they choose for themselves which trees to explore and
	report anything that they deemed interesting. Specifically, given trees rooted
	in the Enron dataset and at least 20 to thirty minutes, the participants
	situated themselves within the persona of a consultant asked to evaluate a
	company's spreadsheet practice, which could be arranged around questions like
	which functions on which the employees most depended, which sheets presented
	the most anomalous or dangerous designs, and so on. Interpretation of
	"spreadsheet practice" can certainly vary from participant to participant, but
	since we wanted a range of views and backgrounds despite the persona, we
	readily accepted this interpretive wiggle room. \par We clarified, furthermore,
	that these notes could be directed either at the quality of the data conveyed
	by the tool or at the tool's conveyance itself. Since the tool's philosophy was
	based in exploration, we decided that it would not have been appropriate to
	restrict what people were to look for, letting them find interesting stories of
	formula construction for themselves. Additionally, either kind of comment -- on
	data or on tool -- would direct us back to the central evaluation; a user's
	silence, or lack of any interesting findings, would tacitly indicate that
	perhaps the tool does not convey information well enough. \par Still, several
	limitations cropped up during the study that complicate the evaluation. For one
	thing, though the source of the data was not explicitly stated at the outset of
	the session, several users studied the file names (unchanged as they were) to
	deduce its roots in finance, if not Enron itself. Because of these notorious
	connotations, then, some were inclined from this moment of discovery to orient
	their exploration around finance-related functions, potentially limiting their
	findings or perceived users for the tool -- though some also said that, without
	this essential context, they could not have properly explored a dataset in the
	first place. Comments like these point out the trade-offs of unguided
	exploration, too: though it might not restrict them to a single purpose, it
	might, at the same time, not afford them any mindset at all with which to
	interpret the data. Another potential problem is that the data which users
	explored was processed and created before we had fully completed the tool,
	meaning some of the counts and patterns in the trees were inaccurate. However,
	we decided that this was not a pressing concern, being that we were primarily
	with whether any interesting reports could be made at all, not with whether
	these reports were 100\% accurate assessments of reality. \par Nevertheless,
	these user studies were a formative moment in evaluating the tool's
	performance, generating a number of directions or corrections in the process.
	We present an overview of responses below, with the four participants recorded
	simply as P1 through P4. \par \subsection{Positive Responses} The inclusion of
	formula examples, grounded in and linking back to the originating spreadsheets,
	proved to be an especial draw to the tool. After all, none of the participants
	professed themselves to be experts with Excel, their self-reported
	proficiencies ranging from "fairly familiar" (P2) to "not extremely familiar"
	(P4); even if they were experts, it would likely be unreasonable to expect an
	evenly distributed familiarity with the 100+ function trees available.
	Nevertheless, the list of actual formulas accessible through a node's
	double-click allowed the users a way to find examples of formulas with exactly
	the configuration they wanted. For example, as P1 explored the VLOOKUP
	function, even though they knew the official API more verbosely explained its

>>>>

	\section{Introduction} Business and research alike owe a lot to spreadsheets, the
	tabular interface that offers users a structured way to store and manipulate
	potentially huge bodies of data. Their allure is in their versatility: while
	the novice can still work without a deep knowledge of programming, they can
	learn to expedite their work by building programs out of several available
	functions \cite{nardi1990spreadsheet}. As such, it should not be surprising
	that when Scaffidi and colleagues estimated that over 50 million U.S. workers
	could be using them by 2012, with 25 million of them writing programs
	\cite{scaffidi2005estimating}. \par

	Considering this ubiquity, it's crucial to understand how people are actually
	using spreadsheets. After all, failures in their practice can be ruinous, as
	demonstrated in EuSpRIG's many horror stories of spreadsheets gone wrong. In
	2013, for example, they detail the influence of an economics paper on debt and
	national growth which shaped many political initiates around it, only to
	discover that the spreadsheets used in the data analysis contained an error
	that reversed the findings. The other stories, likewise, do not augur well if
	shoddy spreadsheet practices persist
	\footnote{http://www.eusprig.org/horror-stories.htm}. \par

	Fortunately, many spreadsheet researchers have assembled and released a number
	of spreadsheet corpora to inform work of how people actually use these tools.
	Collections like EUSES \cite{fisher2005euses}, FUSE \cite{barik2015fuse}, and
	the Enron corpus \cite{hermans2015enron} have already enabled fruitful work
	across the field. Hermans and colleagues, for example, used the EUSES corpus in
	2011 to inform and evaluate their work on exploring code smells, or inelegant
	places in code that suggest the need for refactoring, between worksheets
	\cite{hermans2012detecting}. Furthermore, these corpora have been compared
	and cross-referenced, such as in Jansen's contrast on Enron and EUSES, to
	gather insight in how different domains use spreadsheets differently 
	\cite{jansen2015enron}, or in Aivaloglou and colleague's evaluation of a
	spreadsheet grammar \cite{aivaloglou2015grammar}.\par

	A tool, then, that empowers users of any intention to explore these datasets
	would be rife with potential. On a simple level, the tool could provide
	concrete statistics on which functions are used in practice most (or least)
	often for a given dataset. Beyond this, though, it could also inform future
	work on recommending functions through an understanding of this patterns, or it
	could guide API improvement by suggesting functions to add or prune based how
	people actually use them. Furthermore, if the tool maintains the connection
	between its broad patterns and the actual instances of formulae in workbooks,
	it could serve as a boon to educators not only by showing which functions and
	combinations are most practical to learn but also by offering a bounty of
	instructive, and real, examples \par

	This paper contributes a tool, informed by some of the aforementioned
	spreadsheet corpora, that seeks to accomplish these tasks by offering an
	interactive interface for exploring how people combine functions. It seeks to
	balance several needs, such as capturing these patterns in a broad enough
	overview while still be specific enough lead back to the precise places in
	spreadsheets where these patterns are realized. Along with the tool, then, is
	also a discussion of the various goals which the tool seeks to fulfill and the
	trade-offs that we negotiated at each point. All of this culminates in an
	evaluation of the tool through a study of possible cases to which the tool
	might apply.

	\section{Related Work} This tool comes from a line of spreadsheet visualization
	tools before it, each with a different focus. Some tools, such as Igarashi and
	colleague's fluid visualizations \cite{igarashi1998fluid}, seek to visualize
	the hidden formulas within a spreadsheet by imposing the dataflow graphs over
	the cells. Likewise, Clermont \cite{clermont2003scalable} (and Hipfl
	\cite{hipfl2008using}, who extended his work) explored various ways of
	visualizing groups of related cells through similar functions, neighbors, or
	references. Others focus on creating visualizations outside of the sheets:
	Hermans and colleagues address a spreadsheet programmer's information needs
	through their work on the tools GyroSAT \cite{hermans2011supporting} and Breviz
	\cite{hermans2011breviz} to make dataflow diagrams of individual spreadsheets.
	These, however, tend to focus on individual spreadsheets and not bodies of
	them. \par

	Nevertheless, other studies and tools focus more on evaluating the content of
	the cells therein. Badame, for example, implements a number of Excel formula
	refactorings and uses Euses in part to evaluate it
	\cite{badame2012refactoring}. Hermans, leveraging her work with the
	aforementioned dataflow diagrams, approached the same dataset to assess the
	prevalence of common inter-worksheet smells, such as feature envy and
	inappropriate intimacy \cite{hermans2012detecting}. Cunha and colleagues,
	likewise, constructed and refined a catalogue of smells and implemented the
	SmellSheet Detective tool to detect bad smells in chosen spreadsheets
	\cite{cunha2012towards}, which Abreu and colleagues later extended to detect
	faults. \cite{abreu2014smelling}.\par

	Implicit in this study is an exploration of how users interact with the Excel
	API. Previous research, such as the first study on the Enron spreadsheet corpus
	\cite{hermans2015enron}, on how users employ certain functions or commands.
	Researchers have applied such questions to other contexts as well, such as
	Murphy and colleague's study into how Java developers use Eclipse and what
	commands they use most often \cite{murphy2006java}.  Others focus on the
	documentation of the API itself: Robillard and DeLine, for example conducted a
	series of surveys and interviews to diagnose common problems with learning
	APIs, and they found that code examples, one of the focuses of this tool, is
	one of the most important aids to have in learning a new system
	\cite{robillard2011field}.

	\section{Approach/Methodology} \subsection{Walkthrough of Tool} Before
	discussing the minutiae of how the tool was developed, it will help first to
	give a basic example of how the tool can be used and how to interpret its
	symbols. One such example might be to find the answer to this question:
	\textit{ What kinds of functions do people use to define the condition in an IF
		function?} The final visualization is available in figure \ref{fig:fullpic};
	however, to understand the interaction, we will start rather from the
	beginning. \par

	When the user first approaches the tree for a given top-level function -- that
	is, a function nested within none other -- only a few nodes are visible, as
	shown in figure \ref{fig:startpic}. \par

@@

	The visualization so far comprises two types of node: the circle, which
	represents a discrete function in the formula and is size according to its
	relative frequency (the top-level node, by definition, will be the largest);
	and the numbered squares, which represent the positions of arguments within its
	parent function. From this, we can infer that, of the times it was observed, IF
	can have at most three arguments passed into it, which corresponds with its
	specifications in the
	API.\footnote{https://support.office.com/en-us/article/IF-function-69aed7c9-4e8a-4755-a9bc-aa8bbff73be2} \par

	Knowing that it is the first argument which contains the conditionals, we click
	on the square labeled "1" to explore. To save space, when more than 10 unique
	arguments have been observed in any position, the tool displays only the first
	ten, with an option to display the rest. The results are shown below:

@@

	As expected, we see that IF contains as its first argument a number of
	comparison operators, such as = and <=, and boolean-returning functions, like
	ISNA and AND, with simple equality being the most common and some use as ABS
	being the least seen of everything actually used. \par

	From here, we can further explore the common options among these functions.
	Clicking on the "=" node will yield two arguments, it being a binary operator,
	and expanding each of them will peer into the range of common values of
	equality comparison, which Figure \ref{fig:fullpic} shows. \par

	For both sides of the equal sign, the operator has certain types of arguments
	which predominate over the others: on the left side is most often a reference
	to another cell; on the right, a string or number literal, which makes sense
	for the case of confirming a value in another cell before assigning this one.
	Furthermore, if the concept is difficult to imagine is practice, a tooltip
	accompanies each function node in the tree, providing a concrete example of a
	function that uses this structure and where it can be found. If this single
	instance is not enough, then the user also has the option to double-click the
	individual function node, which will open up a new tab with a table of many
	more examples from dataset.

@@

	\subsection{Design Goals} In visualizing the spreadsheet data, I outlined a few
	core goals for what the tool should accomplish: \begin{itemize} \item [1]
		\textbf{Draw an interactive interface to explore observed function
			combinations.} The combination space for all Excel functions is massive, let
		alone the space for observed formulas. Working from data with actual referents
		in practice, the tool must aid the user in navigating this space. \item [2]
		\textbf{Emphasize the quantitative patterns in formula construction.} The tool,
		accommodating datasets of a few spreadsheets to millions, should address the
		questions of how often the end users employed a certain function and where they
		used it. In this way, it must show precise metrics, such as frequency of  use
		and depth of function nesting, of the dataset it conveys. \item [3]
		\textbf{Promote a qualitative understanding of the patterns.} Lest these
		patterns remain abstract, the tool should supplement its observations with
		concrete instances of relevant formulas from the corpus. Ideally, it should
		even direct users to the exact cell in the spreadsheets where the formula was
		used, contextualizing the functions. \end{itemize} \par

	Likewise, to bound our scope, we outlined a few goals to specifically avoid
	accomplishing with this: \begin{itemize} \item [!1] \textbf{Do not try to
			directly explain what a function does.} Though the tool tries to foster tries
		understanding by linking pattern to example, it won't provide a precise
		description of what a function accomplishes. The tool's user must infer this.
		\item [!2] \textbf{Do not create new formulas.} This is essentially an
		exploratory tool, not a generative one. It should not produce any information
		other than new views of the original data. \item [!3] \textbf{Do not visualize
			individual formulas.} Though there is room to explore the place of a single
		function in the group, the tool must design the core visualization around the
		entire body of data, not the other way around. \end{itemize}

	\subsection{Design Decisions} Early in the process, we decided that the tree
	form would be a suitable fit to represent the parent-child or caller-callee
	relationships inherent in the data, given the composition of formulas as
	functions and their arguments (which could be yet more functions). By adapting
	this structure to accommodate a broad range of possibilities for nested
	functions, the branching factor depends on both the number of arguments in a
	function and the number of possible functions observed as an argument in a
	function. \par

	Guided by the goals in section A, we faced a number of decision points, a
	sample of which we describe below, before we arrived at the design shown above.
	\begin{itemize} \item \textit{Copied formulas}: Excel allows users to spread a
		formula over an area, repeating the same task in each cell with minor
		adjustments. Without checking for this, the analysis may not reveal the
		functions most commonly used together but rather the formulas most often
		applied to large areas. To combat this, we converted formulas from their native
		A1 format to the relative R1C1, in which copied formulas should be identical,
		and reduced the records to unique R1C1 formulas per sheet. \item
		\textit{Importance of depth}: When a function appears within another, should it
		be analyzed only as a nested function or would it also be valid to analyze the
		nested function on its own?

		\par For example, in the pictured IF function, we found that people often use
		another IF statement as the second or third arguments of the top-level IF. If
		we only consider functions exactly where they're found embedded in the
		formula, then information about the same function -- IF, in this case -- will
		be scattered across different trees with no way to aggregate them. If we
		record every instance of a function by ignoring their context -- that is,
		including an IF embedded within a SUM function in the same node as the
		top-level IF -- then the tool will represent some functions in multiple nodes
		to capture every possible level of nesting. Both approaches have benefits and
		drawbacks, and so we included both. \item \textit{Pattern density}: How should
		the tool quantitatively order its elements: by a function's raw frequency or
		by the unity of patterns it leads to? For example, if SUM has for its first
		argument two possibilities, one which itself contains 1000 unique argument
		possibilities with 1 occurrence each (high frequency, low pattern density) and
		another seen with 2 argument possibilities of 100 occurrences each (low
		frequency, high pattern density), which would be more interesting to
		emphasize? The answer depends on the nuances of the questions, but for
		simplicity, I've shown the former. \item \textit{Non-functions}: How should
		the tool represent everything in the formula that isn't a function: numbers,
		string literals, errors, references, etc? Since these don't accept arguments,
		they will adorn the tree as leaves, and their precise content won't affect the
		functions around them as long as their types are known. As such, in the
		visualization, all of these nodes are replaced and aggregated under their
		types and represented as empty white nodes. \item \textit{Optional arguments}:
		How should the tool handle functions which accept a variable number of
		arguments? SUM, for example, can have anywhere from 1 to 255, and IF can
		accept either 2 or 3.
		\footnote{https://support.office.com/en-us/article/Excel-functions-by-category-5f91f4e9-7b42-46d2-9bd1-63f26a86c0eb} Without prior evidence, it's possible that, when people use optional arguments, they use different kinds of types and functions for arguments versus the cases when they ignore optional arguments. To account for this, we separate and analyze the different quantities of arguments observed for each function; the tool, however, uses as default all of the options collapsed into a single representation. \end{itemize}

@@

	\subsection{Bad Smells} Fowler's description of code smells \cite{fowler2009refactoring}
	underscored an important point of code quality: between perfect code and
	bug-crippled spaghetti,  there is a spectrum of code designs which, by
	themselves, are not faulty but nevertheless suggest problems and
	vulnerabilities in code design. Code smells, as we call them, soon commanded
	attention and earned wide-ranging applicability, eventually making it to the
	art of spreadsheet design.

	Since then, spreadsheet researchers have taken the concept, if not the specific
	smells themselves, and applied them to spreadsheet programming
	\cite{hermans2012detecting} \cite{cunha2012towards} \cite{jansen2015code}
	\cite{asavametha2012detecting} . While some of the smells lie beyond the 
	scope of the tool, such as those characterizing inter-worksheet connections,
	some have structures which create distinctive features within the tree,
	leading to easy detection through visualization: \par

	\begin{itemize}
		\item \textit{Long Method}: Also called "Multiple Operations" in \cite{hermans2012detecting},
		this refers to situations where numerous functions and operators are used all within the same
		cell, creating a formula that tends to extend beyond a compact viewing space and makes it
		difficult to understand. In the visualization, since each nested and combined function creates
		yet another level of depth within the tree, these constructions can result in long, horizontal 
		chains of nodes, as seen in figure \ref{"fig:longsum"}.
		\item \textit{Long Parameter List}: [tall, vertical towers of blue nodes]
		\item \textit{Conditional Complexity}: [IF nodes within IF nodes]
	\end{itemize}

	\begin{figure*}
		\centering
		\includegraphics[width=\textwidth]{longsum}
		\label{"fig:longsum"}
		\caption{The horizontal length of a tree may indicate complex or redundant design.}
	\end{figure*}

	Some examples are meaningful not for the space they take up but for how they
	clash with the expectations around a function. For example, Excel offers a
	number of functions that accept any number arguments given to them, up to a
	resource-defined limit. SUM and CONCATENATE are two such functions: adding more
	arguments simply adds or appends that element in the same way as every previous
	argument. Strange cases were found, then, in the cases where these functions
	were used with only one argument. \par

	%Lookup example
	Otherwise, the LOOKUP functions are particularly notorious for their confusing
	mechanisms and, as such, have warranted some especial attention with regards to
	spreadsheet design. In a recent study, Hermans, Aivaloglou, and Jansen probed
	the Enron dataset for uses of VLOOKUP and HLOOKUP specifically.
	The issue at hand was the optional fourth argument: when omitted or set to TRUE,
	the function attempts to return a value from another column by matching two values
	approximately, but when set to FALSE, the match must be exact. Furthermore, exact
	matching works only with sorted columns or rows; if the programmer neglects to sort
	values before calling the *LOOKUP function, they may acquire inaccurate results \cite{hermans2015detecting}. \par

	Given access to the same dataset, how might this tool approach the issue? 
	On the surface, the tool supports easy quantification through the tooltips
	displayed for each node, as shown in figure \ref{"fig:vhlookups"}, answering
	the question, "How often do spreadsheet programmers use this function?". But, 
	more to the point, the tool's distinction between the numbers of arguments
	for a single function lets the user compare frequencies of three-argument 
	forms with the four-argument. The visualization does not, however, discern
	between the different values for booleans in the tree, much less whether 
	the spreadsheet columns are ordered -- but the linking of examples to patterns
	allows quick insight into which files are at risk for the problems they outlined. \par

	\begin{figure}
		\centering
		\includegraphics[width=.5\textwidth]{vlookup}
		\includegraphics[width=.225\textwidth]{vlookupbox}
		\includegraphics[width=.225\textwidth]{hlookupbox}
		\includegraphics[width=.5\textwidth]{hlookup}
		\label{"fig:vhlookups"}
		\caption{VLOOKUP and HLOOKUP with four arguments. Numbers
			not expected to correspond perfectly with \cite{hermans2015detecting}.}
	\end{figure}

	Also at hand are issues with the standard LOOKUP function. Unlike the two 
	variations above, LOOKUP accepts either two or three arguments, depending
	on which API-defined form the programmer uses: the vector form or
	the array form, the primary difference being whether the second parameter
	is a one-column/-row range or a two-dimensional array. The array form, as 
	it happens, is strongly discouraged by the API in favor of the V/HLOOKUP,
	a notion that might also be leveraged in the pursuit in good spreadsheet 
	design.

	\begin{figure}
		\centering
		\includegraphics[width=.5\textwidth]{lookup-2}
		\includegraphics[width=.5\textwidth]{lookup-3}
		\label{"fig:lookups"}
		\caption{LOOKUP with two (6 occurrences) and three (1648) arguments.}
	\end{figure}

@@

	\section{User Study} Midway through production, we conducted a brief,
	four-participant user study in order to gauge how well we were capturing the
	initial design goals. The study was, in general, a flexible and exploratory
	affair: rather than giving the users any specific tasks or pointed questions,
	we asked instead that they choose for themselves which trees to explore and
	report anything that they deemed interesting. Specifically, given trees rooted
	in the Enron dataset and at least 20 to thirty minutes, the participants
	situated themselves within the persona of a consultant asked to evaluate a
	company's spreadsheet practice, which could be arranged around questions like
	which functions on which the employees most depended, which sheets presented
	the most anomalous or dangerous designs, and so on. Interpretation of
	"spreadsheet practice" can certainly vary from participant to participant, but
	since we wanted a range of views and backgrounds despite the persona, we
	readily accepted this interpretive wiggle room. \par

	We clarified, furthermore, that these notes could be directed either at the
	quality of the data conveyed by the tool or at the tool's conveyance itself.
	Since the tool's philosophy was based in exploration, we decided that it would
	not have been appropriate to restrict what people were to look for, letting
	them find interesting stories of formula construction for themselves.
	Additionally, either kind of comment -- on data or on tool -- would direct us
	back to the central evaluation; a user's silence, or lack of any interesting
	findings, would tacitly indicate that perhaps the tool does not convey
	information well enough. \par

	Still, several limitations cropped up during the study that complicate the
	evaluation. For one thing, though the source of the data was not explicitly
	stated at the outset of the session, several users studied the file names
	(unchanged as they were) to deduce its roots in finance, if not Enron itself.
	Because of these notorious connotations, then, some were inclined from this
	moment of discovery to orient their exploration around finance-related
	functions, potentially limiting their findings or perceived users for the tool
	-- though some also said that, without this essential context, they could not
	have properly explored a dataset in the first place. Comments like these point
	out the trade-offs of unguided exploration, too: though it might not restrict
	them to a single purpose, it might, at the same time, not afford them any
	mindset at all with which to interpret the data. Another potential problem is
	that the data which users explored was processed and created before we had
	fully completed the tool, meaning some of the counts and patterns in the trees
	were inaccurate. However, we decided that this was not a pressing concern,
	being that we were primarily with whether any interesting reports could be made
	at all, not with whether these reports were 100\% accurate assessments of
	reality. \par

	Nevertheless, these user studies were a formative moment in evaluating the
	tool's performance, generating a number of directions or corrections in the
	process. We present an overview of responses below, with the four participants
	recorded simply as P1 through P4. \par
